Exploiting Private and Hybrid Clouds for Data-intensive Computing

### 2012-08-03 MSc-dissertation.tex
"Aug"

{\bf Origin of the term Big-data}

With more than 2.5 quintillion bytes of data \cite{website:ibm-bigdata} being produced
every day, every now and then we run into a problem of not having 
sufficient computing resources to process and extract useful information
within the required time-frame.

The term big-data is time specific and broadly
big data can be defined at any point in time as ``data whose size forces us
 to look beyond the tried-and-true methods
that are prevalent at that time.`` In the early 1980s, it was a dataset that was
 so large that a robotic ``tape monkey'' was required to swap thousands of tapes in and out.
In the 1990s, perhaps, it was any data that transcended the bounds of Microsoft
Excel and a desktop PC, requiring serious software on Unix workstations to
analyze. Nowadays, it may mean data 
that is too large to be placed in a relational database and analyzed with the
help of a desktop statistics/visualization package-data, perhaps, whose
analysis requires massively parallel software running on tens, hundreds, or
even thousands of servers.\cite{jacobs2009pathologies}

@@

The shelf-life of the data is short so should be able to act on it before it expires.

Cost effectively process and analyze this database

@@

According to IBM, the challenges of big-data spans three dimensions: Volume,
Velocity and Variety.\cite{website:ibm-bigdata}

Volume:

Enterprises are awash with ever-growing data of all types,
easily amassing terabytese-ven petabytes-of information.

@@

Velocity: Sometimes 2 minutes is too late. For time-sensitive processes such as catching fraud,
big data must be used as it streams into your enterprise in order to maximize its value.

@@

Variety: Big data is any type of data - structured and unstructured data such as text,
sensor data, audio, video, click streams, log files and more. New insights are found when
analyzing these data types together.

Monitor 100’s of live video feeds from surveillance cameras to target points of interest
Exploit the 80\% data growth in images, video and documents to improve customer satisfaction

@@

Apache Hadoop is a software framework for distributed processing of huge data-sets
using the MapReduce programming model. The Apache Hadoop

The open source Apache Hadoop software library is a framework that 
allows the distributed processing of large data sets across clusters of 
computers using the MapReduce programming model \cite{website:apache-hadoop}. It also provides 
a distributed ﬁle system, the Hadoop Distributed File System, that provides 
high-throughput access to application data.

>>>>

% Origin of the term Big-data
% ---------------------------

With more than 2.5 quintillion bytes of data \cite{website:ibm-bigdata} being produced
every day, every now and then we run into a problem of not having 
sufficient computing resources to process and extract useful information
within our required time-frame.

The term big-data is time specific and broadly
big data can be defined at any point in time as ``data whose size forces us
 to look beyond the tried-and-true methods
that are prevalent at that time.`` In the early 1980s, it was a dataset that was
 so large that a robotic ``tape monkey'' was required to swap thousands of tapes in and out.
In the 1990s, perhaps, it was any data that transcended the bounds of Microsoft
Excel and a desktop PC, requiring serious software on Unix workstations to
analyze. Nowadays, it may mean data 
that is too large to be placed in a relational database and analyzed with the
help of a desktop statistics/visualization package-data, perhaps, whose
analysis requires massively parallel software running on tens, hundreds, or
even thousands of servers \cite{jacobs2009pathologies}.

@@

The shelf-life of the data is short so we should be able to process the data cost effectively,
before its shelf-life expires.

@@

According to IBM, the challenges of big-data spans three dimensions: Volume,
Velocity and Variety \cite{website:ibm-bigdata}.

\item \emph{Volume} \\
Enterprises are awash with ever-growing data of all types,
easily amassing terabytese-ven petabytes-of information.

@@

\item \emph{Velocity} \\
Sometimes 2 minutes is too late. For time-sensitive processes such as catching fraud,
big data must be used as it streams into your enterprise in order to maximize its value.

@@

\item \emph{Variety} \\
Big data is any type of data - structured and unstructured data such as text,
sensor data, audio, video, click streams, log files and more. New insights are found when
analyzing these data types together.

Monitor 100's of live video feeds from surveillance cameras to target points of interest
Exploit the 80\% data growth in images, video and documents to improve customer satisfaction

\end{itemize}

@@

Apache Hadoop \cite{website:apache-hadoop} is a top-level Apache project that includes open source implementations of a distributed file system \cite{website:apache-hdfs} and MapReduce that was inspired by Google's GFS
\cite{ghemawat2003google} and MapReduce \cite{dean2008mapreduce} projects.

The open source Apache Hadoop software library is a framework that 
allows the distributed processing of large data sets across computer cluster using the MapReduce programming model \cite{website:apache-hadoop}. It has a distributed file system, the Hadoop Distributed File System, that provides 
high-throughput access to application data.

Facebook uses the Apache Hadoop platform to run it messages application \footnote {Facebook Messages gives every 
Facebook user a facebook.com email address, integrates the display of all e-mail, SMS and chat messages between a pair or group of users, has strong controls over who users receive messages from, and is the foundation of a Social Inbox\cite{borthakur2011apache}.}. Hadoop was chosen because it met the application's reqirements for consistency, availability, partition tolerance, data model and scalability \cite{borthakur2011apache}.

The Hadoop ecosystem also includes projects like:\cite{website:apache-hadoop}

\begin{itemize}
\item \emph{Avro:} A data serialization system.
\item \emph{Cassandra:} A scalable multi-master database with no single points of failure.
\item \emph{Chukwa:} A data collection system for managing large distributed systems.
\item \emph{HBase:} A scalable, distributed database that supports structured data storage for large tables. This was inspired by Google's BigTable.
\item \emph{Hive:} A data warehouse infrastructure built on top of Hadoop that provides data summarization and ad hoc querying.
\item \emph{Mahout:} A Scalable machine learning and data mining library.
\item \emph{Pig:} A high-level data-flow language and execution framework for parallel computation.
\item \emph{ZooKeeper:} A high-performance coordination service for distributed applications.
\end{itemize}
